\chapter{Attunements}

\begin{multicols}{2}

\section{About Attunements}

The entries below are the attunements of Vox Draconis. You may
never have more than one attunement.

An attunement is a supernatural alignment to a particular
constellation or heavenly body. It is rare among mortals,
affecting only 1 in 10,000. Those born with an attunement
are destined for great and sometimes terrible things. Some
animals also have attunements, though that is even rarer than
among sentient beings.

This grants you a supernatural ability and an elemental affinity.
The affinity \index{affinity} makes you immune to damage or
ill effects from that element, but also means you take double
damage or worsened effects from its opposite. The exact nature
depends on the circumstances; for example, if you have an
affinity to fire, and you fail a swimming check while trying to fight
rapids, you may have to make a Crisis Roll to avoid drowning
rather than simply be washed downstream.

\subsection{The Elements}

Air opposes Earth.

Fire opposes Water.

If you have an affinity for Air, you are weak against Earth. So
too is someone with an affinity for Earth weak against Air.

The same applies for Fire and Water.

\section{List of Attunements}

\subsection{Bull}

The Bull is a weighty cluster of white and red stars in
the western sky.

\textbf{Affinity:} Earth

\subsubsection{Ability: Stubborn Charge}

Once per scene, you can charge up to your speed and nothing stops you once you get
moving. You destroy any obstacle you charge into, and any living thing
you charge into takes 1d10 damage and is flung aside or trampled.

\subsection{Dragon}

The constellation Dragon appears high in the sky. Its twisting
shape is memorable, but the most notable thing about it are
the two orange stars that make up its "flame."

\textbf{Affinity:} Fire

\subsubsection{Ability: Breath Fire}

Once per scene, you can breath fire. This can be a small puff of flame
or up to a dog-sized gout of fire. The flame is as hot as a campfire. It
deals 1d4 damage to those caught in it, and may ignite flammable
materials.

\subsection{Moon}

The Moon of Yrda is large and bright. It has long been considered
the ruler of night.

\textbf{Affinity:} Water

\subsubsection{Ability: Silver Glow}

At night, you are surrounded by a faint silver glow. While in this state,
you can see perfectly well even in complete darkness.

\subsection{Phoenix}

The constellation Phoenix spreads large wings. It rises up out of the
deep southern sky.

\textbf{Affinity:} Fire

\subsubsection{Ability: Rebirth}

When you die, you are reincarnated in a newborn. When that person
reaches adolescence, they begin having flashbacks and dreams of their
past lives. When they reach adulthood, they possess all the memories
and knowledge of their past lives, though none of their skills or personality.

\subsection{River}

The constellation River winds its way through the low western sky.

\textbf{Affinity:} Water

\subsubsection{Ability: Waterbreathing}

You can breath as well in water as you can in air.

\subsection{Serpent}

The stars of Serpent stretch in a curving line across the
southern sky.

\textbf{Affinity:} Earth

\subsubsection{Ability: Venomous}

You have unusually sharp incisors, and your bite is venomous. Any
living thing bitten by you suffers 1d4 poison damage every round for
four rounds.

\subsection{Sky}

The Sky is a bright blue blanket enveloping the world.

\textbf{Affinity:} Air

\subsubsection{Ability: Wingless Flight}

Once per scene, you can fly up to a hundred feet. If you are
still in the air once you reach that limit, you fall to the ground.

\subsection{Spire}

The constellation Spire has seven bright green stars arranged
like a spear reaching upward.

\textbf{Affinity:} Earth

\subsubsection{Ability: Displace Earth}

Once per scene, you can cause a section of earth to rise or lower
abruptly, up to six feet. The section can be no larger than ten
feet wide and ten feet across. Anything standing on the ground
must make a Hard Physical Dexterity roll. If they fail, they fall
over as the ground under them shifts violently.

\subsection{Spirit}

The Spirit constellation is ghostly, possessing three pale nebulae amongst
its stars.

\textbf{Affinity:} Water

\subsubsection{Ability: Become Wraithlike}

Once per day, you can become incorporeal for up to a scene. In this state,
you cannot be seen, heard, or touched. You can pass through solid objects
at will, but you cannot affect the physical world. You can hear and see
everything around you.

\subsection{Storm}

Storms on Yrda are short-lived but ferocious. So, too, are those
with this attunement.

\textbf{Affinity:} Air

\subsubsection{Ability: Tempest}

Whenever you are angry, lightning appears in your eyes, and clouds
gather above. Once per day, this can grow into a full storm that lasts
for the rest of the scene. Lightning strikes nearby at random, rain pours
down, and wind howls at gale force.

\subsection{Sun}

Yrda's bright orange-yellow Sun brings life and light.

\textbf{Affinity:} Fire

\subsubsection{Ability: Illumination}

Once per day, you can glow so brightly that it appears to be daylight
for a mile in every direction. Attempting to look at you is like looking
directly at the sun. This lasts for up to a single scene.

\subsection{Void}

The Void of the night sky between the stars is an empty expanse.

\textbf{Affinity:} Air

\subsubsection{Ability: Extinguish}

You can immediately extinguish any fire of any size that you can see. This
does not affect stellar bodies.

\end{multicols}
