\chapter{Enemies and Allies}

\begin{multicols}{2}

\section{About Enemies and Allies}

These are the non-player characters players might encounter in 
their adventures. Each is presented with a different set of 
statistics than player characters. When acting, they use their 
own abilities as described here rather than the rules for players. 
Where the rules are unclear for a situation, the Game Master decides 
how to proceed.

Each being in this section is described with the following format.

\begin{itemize}
    \item \textbf{Name:} the name of the being
    \item \textbf{Desires:} what drives the being to act
    \item \textbf{Fears:} what compels the being to attack or flee
    \item \textbf{Description:} a description of what the being looks like, how they move, and how they act
    \item \textbf{Abilities:} statistics and rules for each of the being's key abilities
\end{itemize}

\section{Dinosaurs}

These are free-roaming, large reptilian creatures. They
are not generally domesticated and have varying levels
of intelligence. Most are social creatures and operate in
groups.

\subsection{Allosaur}

\subsection{Ankylosaur}

\subsection{Brontosaur}

\subsection{Dilophosaur}

\subsection{Raptor, Greater}

\subsection{Tyrannosaur}

\subsection{Velociraptor}

\section{City Dwellers}

These people live in or near cities.

\subsection{Bandit}

\subsection{City Guard}

\subsection{Cultist}

\subsection{Noble}

\subsection{Ranger}

\section{Elder Beings}

These beings are sentient and have lived on Yrda for much
longer than the smaller species that inhabit cities.

\subsection{Dragon, Great Horned}

\subsection{Dragon, Lake}

\subsection{Dragon, Moon}

\subsection{Dragon, Rock}

\subsection{Dragon, Forest}

\section{Wild Folk}

These are sentient beings that live in the wilderness alone or
in very small groups.

\subsection{Ogre}

\subsection{Troll}

\end{multicols}
