\chapter{The Game Setting}

\begin{multicols}{2}

\section{About the World of Yrda}

Yrda is a strange and wondrous place. There are lush jungles full
of dangerous beasts and dinosaurs. There are expansive deserts with
ancient ruins and powerful artifacts hidden beneath the dunes. There
are long mountain ranges punctuated by violent, active volcanos.
All of these places are inhabited to one degree or another. Civilizations
of varying degrees of technological progress and wildly different
cultures exist all over Yrda.

Chief among these are the authoritarian Kingdom of Makhet, the
spiritual Inyani, the imperialistic Kingdom of Gaddar, and the
mysterious Illdrazi. Not all are political, but all are important
in their own right.

All of these lay on the continent of Bardu - the known world at this
time. What lies beyond the vast oceans is unknown.

\section{Anima}

The world of Yrda is suffused with a mystical energy called \textit{anima}. \index{anima}
This energy is the source of most magic. Anima is drawn to places and creatures
that use it, and its nature shifts over time to reflect how it is used.

For example, a place where a lot of healing magic is used might have a
calming, soothing anima. A place where a lot of destructive magic is used
might have a violent, chaotic anima. Areas with a significant elemental
focus might have anima that is associated with that element.

\section{Important Kingdoms and Cultures}

\subsection{The Kingdom of Makhet}

Authoritarian kingdom with a heavily oppressed populace. The rule
of law is considered more important than anything else.

NOTE: Makhet is based on Babylon.

\subsection{The Kingdom of Gaddar}

NOTE: Gaddar is based on the Minoans.

\subsection{The Empire of Sushani}

NOTE: Sushani is based on the Mongols.

\subsection{The Illdrazi}

A culture built around dragons and the mystical bond between the two.

\subsection{The Ardonan Reaches}

Steamy jungles, home to tribal cultures that live amongst dinosaurs.

\subsection{The Inyani}

This culture is built around ancester worship, holy gardens, and the
good of the community.

\section{Important Places}

\subsection{The Wild Tooth Mountains}

A series of treacherous mountains with a handful of passes allowing
travel through them. There is more than one secret place hidden here.
One of those is the Den of Purity, a valley that only Soul Hunters
can find.

\section{The Great Powers}

\subsection{The Unknowable Void}

A formless and vast power that can consume even the gods themselves. It
is not sentient, but hungers always.

\subsection{The Fell Titan}

Somewhere between a god and a demon, the Fell Titan is an incorporeal
entity that fights with the other Great Powers for influence over the
mortal world. It is most closely associated with the bridge between life
and death.

\subsection{The Path of Three}

The Light, the Dark, and the Balance form the trinity of this pantheon.

\section{Languages}

The following are some of the languages in Yrda:

\begin{itemize}
  \item Yrdish - the common language of Yrda
  \item Old Yrdish - a more archaic version of Yrdish, this is no longer spoken
    outside of scholarly circles or some lost cultures
  \item Gaddari - the language of the Gaddar people
  \item Makhetian - the language of the Makhet culture, spoken primarily in the
    Kingdom of Makhet
  \item Ardonan - the language of the Ardonan culture, spoken primarily in the
    Ardonan Reaches and western parts of the Sushani Empire
  \item Sushani - the language of the Sushani culture, spoken primarily in the
    Sushani Empire
\end{itemize}

There are many lands and cultures in Yrda. The Game Master and the players can work
together to make up new languages.

\end{multicols}
